Improving the Accuracy of a Sequential Mining Model for Hurricane Trajectory Prediction

### 2018-02-05 paper/manuscript_revised.tex
Rename manuscript to remove specific date

\maketitle
% Clearly state that weighted support is our idea, inspired by some work but we have a diff approach
\begin{abstract}

@@

Consider figure (2), generated using a sample of the recorded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}. As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape. A discretization block size of 5 is selected in HTPDM and is replicated in WARD-HTP. A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km. The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudinal points \cite{lat-long-distance}. The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.

@@

In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the distance in kilometres between two latitude-longitude points over a large sphere, such as the earth's surface. Despite the fact that the earth's surface is not a perfect sphere \cite{lat-long-distance}, the Haversine Distance Formula is sufficiently accurate when used within a specific region such as the Atlantic Basin. The two formulas, denoted $D_{eucl}$ and $D_{havr}$ are defined as follows:
\begin{align*}
& (1)\ D_{eucl}(p, q) = \sqrt{(\Delta Lat)^2 + (\Delta Lon)^2}\\
& (2)\ D_{havr}(p, q) = 2R(\atantwo(\sqrt{\alpha}, \sqrt{1-\alpha}))\\
& \alpha = \sin^2(\frac{\Delta Lat}2) + \cos(p_{lat}) + \cos(q_{lat}) + \sin^2(\frac{\Delta Lon}2)
\end{align*}

% H is in RADS, Eucl in degrees
Where $\Delta Lat$ and $\Delta Lon$ respectively denote the differences in latitude and longitude between points q and p. Note that the Euclidean distance formula expresses Latitude and Longitude in degrees and minutes, while the Haversine formula expresses Latitude and Longitude exclusively in radians. In addition, whereas the Euclidean formula expresses distance in terms of units on a plane, the Haversine formula expresses distance in terms of portions of a single unit sphere. Therefore, to express distance in kilometres, the result of Haversine formula must be multiplied by R, which denotes the radius of the earth in kilometres (approximately 6371km \cite{earth-radius-kilometres}).

@@

\section{Conclusions and Future Work} % what we found:

@@

\begin{thebibliography}{1}
\bibitem{hurricane-cost} Pielke Jr, Roger A., et al. Normalized hurricane damage in the United States: 1900–2005. Natural Hazards Review 9.1 (2008): 29-42
\bibitem{hurricane-cost-non-financial} F. F. Townsend, The Federal Response to Hurricane Katrina: Lessons Learned, United States Federal Government, Washington, DC, USA, 2006
\bibitem{data mining-def} S. Sumathi and S. N. Sivanandam, Introduction to Dataming and its Applications: Springer Publishing 2006, ISBN: 978-3-540-34350-9
\bibitem{major-hurricane-model} Dong, X and C. Pi, D. (2013). Novel method for hurricane trajectory prediction based on data mining. Natural Hazards and Earth System Sciences. 13. 3211-3220. 10.5194/nhess-13-3211-2013. 
\bibitem{HURDAT2-original} C. Landsea, J. Franklin and J. Beven (2015) The revised Atlantic hurricane database (HURDAT2). National Oceanic and Atmospheric Administration
\bibitem{AprioriAll-original} R. Agrawal and R. Srikant, Mining sequential patterns, Proceedings of the Eleventh International Conference on Data Engineering, Taipei, 1995, pp. 3-14.
doi: 10.1109/ICDE.1995.380415
\bibitem{hurricane-intensity-1} Y. Su, S. Chelluboina, M. Hahsler and M. H. Dunham, A New Data Mining Model for Hurricane Intensity Prediction, 2010 IEEE International Conference on Data Mining Workshops, Sydney, NSW, 2010, pp. 98-105
doi: 10.1109/ICDMW.2010.158
\bibitem{hurricane-intensity-2} R. Yang, J. Tang and D. Sun (2011), Association Rule Mining Application for Atlantic Tropical Cyclone Intensity Changes, American Meteorological Society. doi: 10.1175/WAF-D-10-05029.1
\bibitem{hurricane-forecasting-1} Rozanova, Olga S., Jui-Ling Yu, and Chin-Kun Hu. Typhoon eye trajectory based on a mathematical model: Comparing with observational data. Nonlinear Analysis: Real World Applications 11.3 (2010): 1847-1861.
\bibitem{hurricane-forecasting-2} Chandan Roy and Rita Kovordanyi, Tropical cyclone track forecasting techniques: A review, 2012, Atmospheric research, (104-105), 40-69. doi: 10.1016/j.atmosres.2011.09.012.
\bibitem{typhoon-clustering} Kim, Hyeong-Seog, et al. Pattern classification of typhoon tracks using the fuzzy c-means clustering method. Journal of Climate 24.2 (2011): 488-508.
\bibitem{matplotlib} J. D. Hunter, Matplotlib: A 2D Graphics Environment, in Computing in Science and Engineering, vol. 9, no. 3, pp. 90-95, May-June 2007.
doi: 10.1109/MCSE.2007.55
\bibitem{lat-long-distance} Menno-Jan, K. (2013). Map use: reading, analysis, interpretation: by Kimberling, AJ, Buckley, AR, Muehrcke, PC and Muehrcke, JO, Redlands, California, ESRI Press Academic, 2012. ISBN 978-1-589-48279-1 Seventh edition.
\bibitem{hurricane-distances} D. Brown. (2013) Forecast Verification: Quantifying Forecast Uncertainty. National Hurricane Conference: Hurricane Readiness for Coastal Communities
\bibitem{earth-radius-kilometres} Chambat, F., and Valette, B. (2001). Mean radius, mass, and inertia for reference Earth models. Physics of the Earth and Planetary Interiors, 124(3), 237-253.
\bibitem{war-mining} W. Wang, J. Yang and P. Yu Efficient  mining of  weighted association rules (WAR), Proc. of the ACM SIGKDD Conf. on Knowledge Discovery and Data Mining, 270-274, 2000.
\bibitem{cai-weighted-thesis} Cai, Chun Hing, Ada Wai-Chee Fu, C. H. Cheng, and W. W. Kwong. Mining association rules with weighted items. In Database Engineering and Applications Symposium, 1998. Proceedings. IDEAS'98. International, pp. 68-77. IEEE, 1998.
\bibitem{hurricane-recording-history} Landsea, C. W., Vecchi, G. A., Bengtsson, L., and Knutson, T. R. (2010). Impact of duration thresholds on Atlantic tropical cyclone counts. Journal of Climate, 23(10), 2508-2519.

>>>>

\maketitle

\begin{abstract}

@@

Consider figure (2), generated using a sample of the recorded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}. As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape. A discretization block size of 5 is selected in HTPDM and is replicated in WARD-HTP. A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km. The block dimension size is approximate because the mapping between coordinate degrees and kilometers is imprecise, and is dependent on location in the case of longitudinal points \cite{lat-long-distance}. The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.

@@

In contrast to the Euclidean formula, the Haversine Distance Formula is designed to determine the distance in kilometers between two latitude-longitude points over a large sphere, such as the earth's surface. Despite the fact that the earth's surface is not a perfect sphere \cite{lat-long-distance}, the Haversine Distance Formula is sufficiently accurate when used within a specific region such as the Atlantic Basin. The two formulas, denoted $D_{eucl}$ and $D_{havr}$ are defined as follows:
\begin{align*}
& (1)\ D_{eucl}(p, q) = \sqrt{(\Delta Lat)^2 + (\Delta Lon)^2}\\
& (2)\ D_{havr}(p, q) = 2R(\atantwo(\sqrt{\alpha}, \sqrt{1-\alpha}))\\
& \alpha = \sin^2(\frac{\Delta Lat}2) + \cos(p_{lat}) + \cos(q_{lat}) + \sin^2(\frac{\Delta Lon}2)
\end{align*}

Where $\Delta Lat$ and $\Delta Lon$ respectively denote the differences in latitude and longitude between points q and p. Note that the Euclidean distance formula expresses Latitude and Longitude in degrees and minutes, while the Haversine formula expresses Latitude and Longitude exclusively in radians. In addition, whereas the Euclidean formula expresses distance in terms of units on a plane, the Haversine formula expresses distance in terms of portions of a single unit sphere. Therefore, to express distance in kilometres, the result of Haversine formula must be multiplied by R, which denotes the radius of the earth in kilometres (approximately 6371km \cite{earth-radius-kilometres}).

@@

\section{Conclusions and Future Work}

@@

\begin{thebibliography}{1}
\bibitem{hurricane-cost} Pielke Jr, Roger A., et al. Normalized hurricane damage in the United States: 1900–2005. Natural Hazards Review 9.1 (2008): 29-42
\bibitem{hurricane-cost-non-financial} F. F. Townsend, The Federal Response to Hurricane Katrina: Lessons Learned, United States Federal Government, Washington, DC, USA, 2006
\bibitem{data mining-def} S. Sumathi and S. N. Sivanandam, Introduction to Data mining and its Applications: Springer Publishing 2006, ISBN: 978-3-540-34350-9
\bibitem{major-hurricane-model} Dong, X and C. Pi, D. (2013). Novel method for hurricane trajectory prediction based on data mining. Natural Hazards and Earth System Sciences. 13. 3211-3220. 10.5194/nhess-13-3211-2013. 
\bibitem{HURDAT2-original} C. Landsea, J. Franklin and J. Beven (2015) The revised Atlantic hurricane database (HURDAT2). National Oceanic and Atmospheric Administration
\bibitem{AprioriAll-original} R. Agrawal and R. Srikant, Mining sequential patterns, Proceedings of the Eleventh International Conference on Data Engineering, Taipei, 1995, pp. 3-14.
doi: 10.1109/ICDE.1995.380415
\bibitem{hurricane-intensity-1} Y. Su, S. Chelluboina, M. Hahsler and M. H. Dunham, A New Data Mining Model for Hurricane Intensity Prediction, 2010 IEEE International Conference on Data Mining Workshops, Sydney, NSW, 2010, pp. 98-105
doi: 10.1109/ICDMW.2010.158
\bibitem{hurricane-intensity-2} R. Yang, J. Tang and D. Sun (2011), Association Rule Mining Application for Atlantic Tropical Cyclone Intensity Changes, American Meteorological Society. doi: 10.1175/WAF-D-10-05029.1
\bibitem{hurricane-forecasting-1} Rozanova, Olga S., Jui-Ling Yu, and Chin-Kun Hu. Typhoon eye trajectory based on a mathematical model: Comparing with observational data. Nonlinear Analysis: Real World Applications 11.3 (2010): 1847-1861.
\bibitem{hurricane-forecasting-2} Chandan Roy and Rita Kovordanyi, Tropical cyclone track forecasting techniques: A review, 2012, Atmospheric research, (104-105), 40-69. doi: 10.1016/j.atmosres.2011.09.012.
\bibitem{typhoon-clustering} Kim, Hyeong-Seog, et al. Pattern classification of typhoon tracks using the fuzzy c-means clustering method. Journal of Climate 24.2 (2011): 488-508.
\bibitem{matplotlib} J. D. Hunter, Matplotlib: A 2D Graphics Environment, in Computing in Science and Engineering, vol. 9, no. 3, pp. 90-95, May-June 2007.
doi: 10.1109/MCSE.2007.55
\bibitem{lat-long-distance} Menno-Jan, K. (2013). Map use: reading, analysis, interpretation: by Kimberling, AJ, Buckley, AR, Muehrcke, PC and Muehrcke, JO, Redlands, California, ESRI Press Academic, 2012. ISBN 978-1-589-48279-1 Seventh edition.
\bibitem{hurricane-distances} D. Brown. (2013) Forecast Verification: Quantifying Forecast Uncertainty. National Hurricane Conference: Hurricane Readiness for Coastal Communities
\bibitem{earth-radius-kilometres} Chambat, F., and Valette, B. (2001). Mean radius, mass, and inertia for reference Earth models. Physics of the Earth and Planetary Interiors, 124(3), 237-253.
\bibitem{war-mining} W. Wang, J. Yang and P. Yu Efficient  mining of  weighted association rules (WAR), Proc. of the ACM SIGKDD Conf. on Knowledge Discovery and Data Mining, 270-274, 2000.
\bibitem{cai-weighted-thesis} Cai, Chun Hing, Ada Wai-Chee Fu, C. H. Cheng, and W. W. Kwong. Mining association rules with weighted items. In Database Engineering and Applications Symposium, 1998. Proceedings. IDEAS'98. International, pp. 68-77. IEEE, 1998.
\bibitem{hurricane-recording-history} Landsea, C. W., Vecchi, G. A., Bengtsson, L., and Knutson, T. R. (2010). Impact of duration thresholds on Atlantic tropical cyclone counts. Journal of Climate, 23(10), 2508-2519.
